### 2014-06-16 Equipment/Equipment_Guide/Equipment_Guide.tex
Equipment: Delete and reorganize old water pressure sensor info

>>>>

\chapter{Water Flow Sensor}
\label{chap:Water_Flow_Sensor}

Arduino Config, Setup, and Data Logging

The Arduinos record voltage data locally, and they transmit voltage data to the host computer via the AMQP messaging protocol. The host computer uses RabbitMQ to act as a broker. Python scripts and the pika module send and receive AMQP messages. The following steps setup the host computer and Arduinos for this purpose.
Server (Host Computer)

Note: the following assumes that the server/host computer’s address is 192.168.1.100. It’s best to configure the host computer with a static IP address using static DHCP or manually.

=== Host Computer (RabbitMQ) ===

1) Install RabbitMQ for your platform and start the RabbitMQ server (./sbin/rabbitmq-server on Mac or Linux).

https://www.rabbitmq.com/

2) Copy the receive\_data.py script from the NIST-FIRE repository to the data logging computer.

3) Install Python (The Anaconda Python distribution from Continuum Analytics is recommended)

4) Install the pika module for Python using

\begin{verbatim}
pip install pika
\end{verbatim}

5) Run the receive\_data.py script as

\begin{verbatim}
python receive\_data.py 192.168.1.100 output.csv
\end{verbatim}

== Host Computer (NTP Time Server) ==

On Windows, download and install the NTP server

http://www.meinbergglobal.com/english/sw/ntp.htm

Mac and Linux use NTP natively.

=== Client (Arduino) ===

== Install pressure\_sensor Sketch on the Arduino ==

1) Install Arduino IDE

http://arduino.cc/

2) Plug the Arduino Yun into the USB port of the host computer.

3) Set the Tools > Board to Arduino Yun and the Tools > Port to the /dev/ttyusbmodemXXXXXX setting that matches the plugged-in Arduino.

4) A Sketch is a script that runs on the Arduino microcontroller. Using the Arduino IDE, install the pressure\_sensor Sketch on the Arduino from the NIST-FIRE repository. This enables the reading and writing of data from the analog-to-digital converter via the REST API on the Arduino. Note: this sketch requires the Adafruit ADS 1015 library from https://github.com/adafruit/Adafruit\_ADS1X15.

== Configure the Arduino ==

1) Unplug the Arduino from the computer, and power on the Arduino from the battery.

2) Connect to the Arduino WiFi network. Access the Arduino web interface (http://arduino.local) with the default password of “arduino”. Set the Arduino name (referred to as <arduino\_name> in the following steps), time zone, and WiFi network information on the Arduino web config. Also, set the “REST API Access” to be “Open”.

3) SSH to the Arduino at root@<arduino\_name>.local and install the SFTP server, ntpclient, pip, and pika (for AMQP) on the Arduino

\begin{verbatim}
opkg update
opkg install openssh-sftp-server ntpclient distribute python-openssl
easy\_install pip
pip install pika
\end{verbatim}

4) By default, the microSD card on the Arduino is mounted at /mnt/sda1. Copy the send\_data.py script from the NIST-FIRE repository to /mnt/sda1 on the Arduino.

5) Go to System > Startup > Local Startup in the Arduino web config and add a startup line to the Local Config section (/etc/rc.local)

\begin{verbatim}
python /mnt/sda1/send\_data.py 192.168.1.100 <arduino\_name> /mnt/sda1/output.csv
\end{verbatim}

6) The default NTP service is not very robust, so we will replace it with ntpclient and have it sync with the NTP server every 2 minutes. Disable the default NTP service using

\begin{verbatim}
/etc/init.d/sysntpd disable
\end{verbatim}

Configure the ntpclient service by editing /etc/config/ntpclient to be

\begin{verbatim}
config ntpserver
        option hostname '192.168.1.100'
        option port     '123'

config ntpserver
        option hostname '0.pool.ntp.org'
        option port     '123'

config ntpserver
        option hostname '1.pool.ntp.org'
        option port     '123'

config ntpserver
        option hostname '2.pool.ntp.org'
        option port     '123'

config ntpserver
        option hostname '3.pool.ntp.org'
        option port     '123'

config ntpdrift                                 
        option freq     '0'

config ntpclient                                
        option interval 120
\end{verbatim}

7) Reset power to the Arduino, and then it should start broadcasting data values within about 2 minutes of booting.

\section{Parts}

Mouser \#: 72-T93YB-1K 
Mfr. \#: T93YB102KT20
Desc.:  Trimmer Resistors - Through Hole 3/8" SQ V/ADJ 1K
\$1.48

Mouser \#:   595-TLV2772IP   
Mfr. \#: TLV2772IP
Desc.:  Operational Amplifiers - Op Amps Dual 2.7V R/R
\$2.36

Mouser \#:   517-929834-02-36-RK 
Mfr. \#: 929834-02-36-RK
Desc.: Headers \& Wire Housings FULL STICK HDR/36POS
\$1.13

Mouser \#:   932-MIKROE-513  
Mfr. \#: MIKROE-513
Desc.:  Jumper Wires WIRE JUMPER MALE TO MALE 15CM 10PK
\$3.00

Mouser \#:   932-MIKROE-197  
Mfr. \#: MIKROE-197
Desc.:  Daughter Cards \& OEM Boards SMARTPROTO PROTO ADAPTER BOARD
\$4.90

Mouser \#:   652-PDB12-H4251102BF    
Mfr. \#: PDB12-H4251-102BF
Desc.:  Potentiometers 1K LINEAR 12MM
\$0.649

\section{Flow Meters}

Fire Research Corp.
http://www.fireresearch.com/product.php?id=fta400

Elkhart Brass Hose and Nozzle Accessories
http://www.elkhartbrass.com/products/accessories/hose-nozzle/multimedia
http://www.elkhartbrass.com/files/aa/downloads/catalog/catalog-f2-02-01.pdf

Akron Brass Zero Torque Nozzle
http://www.akronbrass.com/1-1-2-x-1-1-2-zero-torque-shutoff

Omega Pressure Transducers
http://www.omega.com/prodinfo/pressuretransducers.html

Omega Flow Meters
http://www.omega.com/pptst/FP7002.html
http://www.omega.com/pptst/FPB100.html

Task Force Tips Sho-Flow
http://www.tft.com/productsearch/product\_details.cfm?product=SF-BF-BF-11L\&ProductGroup=Sho-Flow\%20Indicator

Flow Equipment
http://www.firehydrant.org/links/flowequip.html

Hose Monster Equipment
http://www.hosemonster.com/products-pages-30.php

Pollard Water
http://www.pollardwater.com/pages\_index/index\_flow\_testing.asp
